Constraining gravitational wave amplitude birefringence with GWTC-3

### 2022-11-26 src/tex/ms.tex
Update ms.tex

% Previous studies
Previous studies have constrained amplitude birefringence by performing Bayesian
inference\kw{This sentence looks odd to me}. In \citet{Okounkova_2022}, they
considered the distribution of observed inclination of the GW events in GWTC-2,
the second GW transient catalog, and use the distribution to constrain the
amplitude birefringence effect with the assumption of frequency independence in
GW amplitude birefringence. In \citet{Yamada_2020} and \citet{Wang_2021}, they
both modeled the waveform under birefringence effect with a simplify assumption
\kw{What is the simplify assumption} and performed Bayesian inference on the
events in GWTC-1, the first GW transient catalog.

% What's new?
In this paper, we used a birefringence model with higher order terms which
include the frequency dependence of amplitude birefringence to perform a more
accurate parameter estimation (PE) \kw{I think there is qualitative difference
between this work and former works, so it is not just more accurate}. We also
performed Bayesian inference on events in GWTC-3, the third GW transient
catalog. We included 69 binary black hole merger events with smallest false
alarm rate $\leq1\mathrm{yr^{-1}}$. These events are also listed in TABLE I in
\citet{GWTC_3_population}.\footnote{GW190720 and GW200129 are included in the
list in \citet{GWTC_3_population}, but we did not include them in this study.
Details in Discussion.}

@@

\section{Method}
\label{sec:Method}
% Waveform modification
GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to
electromagnetic waves, which could be transformed into two circular
polarizations (i.e. left-handed and right-handed) by
\begin{equation}
    h_{\mathrm{L/R}} = \frac{h_+ \pm i h_\times}{\sqrt{2}}\,,
\end{equation}
where $h_{\mathrm{L/R}}$ are the Fourier amplitude of the left-handed and
right-handed polarization of the waveform, $h_+$ and $h_\times$ are the Fourier
amplitude of the plus and cross polarization of the waveform. According to
\textbf{[modification reference]}, we modified the waveform model by
\begin{equation}
    h_\mathrm{L/R}^{\mathrm{br}}=
    h_\mathrm{L/R}^{\mathrm{GR}}\times
    \exp\left(\pm\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}\right)\,,
\end{equation}
where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the modified
waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the
Fourier amplitude of the waveform at the source, $\kappa$ is the dimensionless
opacity parameter that represent the strength of the birefringence, $d_C$ is the
comoving distance to the source, and $f$ is the Fourier frequency of the
waveform. We assume GWs are generated by the binary as GR predicts, as the
modification at the source is small before the effect accumulate during the
propagation. As the GWs propagate, the effect of birefringence will be built up
with the number of cycles which depends on the distance traveled and the
frequency of the GWs.

% Effect on inclination PE
This modification changes the ratio of the left-handed and right-handed polarization of the waveform, which affects the PE on the inclination as well.
In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for a nonprecessing binary.
The relationship between the amplitude ratio and the inclination is
\begin{equation}
    \frac{h_\mathrm{L}}{h_\mathrm{R}}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,,
\end{equation}
where $h_\mathrm{L}$ and $h_\mathrm{R}$ are the Fourier amplitude of left-handed and right-handed polarizations of the GWs and $\iota$ is the inclination, the angle between our line of sight and the orbital angular momentum of the binary.
PE on $\iota$ takes the amplitude ratio into account. Therefore, the PE on $\iota$ depends on the amplitude ratio.

@@

For some previous studies on the birefringence property such as \citet{Okounkova_2022}, they assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
This would create a degeneracy between $\kappa$ and $\iota$, as they can affect the amplitude ratio in the same way.
To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
As a result, many pairs of $\iota$ and $\kappa$ can be plausible with the frequency independence.

>>>>

% Previous studies
Previous studies have constrained amplitude birefringence by performing different statistical analysis.
In \citet{Okounkova_2022}, they considered the distribution of observed inclination of the GW events in GWTC-2,
the second GW transient catalog, and use the distribution to constrain the
amplitude birefringence effect with the assumption of frequency independence in
GW amplitude birefringence. In \citet{Yamada_2020} and \citet{Wang_2021}, they
both modeled the waveform under birefringence effect with a simplify assumption
\kw{What is the simplify assumption} and performed parameter estimation (PE) on the
events in GWTC-1, the first GW transient catalog.

% What's new?
In this paper, we used a birefringence model with higher order terms which
include the frequency dependence of amplitude birefringence to perform a more
accurate PE \kw{I think there is qualitative difference
between this work and former works, so it is not just more accurate}. We also
performed Bayesian inference on events in GWTC-3, the third GW transient
catalog. We included 69 binary black hole merger events with smallest false
alarm rate $\leq1\mathrm{yr^{-1}}$. These events are also listed in TABLE I in
\citet{GWTC_3_population}.\footnote{GW190720 and GW200129 are included in the
list in \citet{GWTC_3_population}, but we did not include them in this study.
Details in Discussion.}

@@

\section{Method}
\label{sec:Method}
% Waveform modification
GW consist of two independent polarization modes similar to electromagnetic waves,
which we can choose to represent them in either the linear basis (i.e. $+$ and $\times$)
or the circular basis (i.e. left-handed and right-handed).
They are related by
\begin{equation}
    h_{\mathrm{L/R}} = \frac{h_+ \pm i h_\times}{\sqrt{2}}\,,
\end{equation}
where $h_{\mathrm{L/R}}$ are the amplitude of the left-handed and right-handed
polarization of the waveform, $h_+$ and $h_\times$ are the Fourier amplitude of
the plus and cross polarization of the waveform. According to
\textbf{[modification reference]}, we modified the waveform model by
\begin{equation}
    h_\mathrm{L/R}^{\mathrm{br}}=
    h_\mathrm{L/R}^{\mathrm{GR}}\times
    \exp\left(\pm\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}\right)\,,
\end{equation}
where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the modified
waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the
Fourier amplitude of the waveform as GR predicts, $\kappa$ is the dimensionless opacity parameter that
represent the strength of the birefringence, $d_C$ is the comoving distance to
the source, and $f$ is the Fourier frequency of the waveform. We assume GWs are
generated by the binary as GR predicts, as the modification at the source is
small before the effect accumulate during the propagation. As the GWs propagate,
the effect of birefringence will be built up with the number of cycles which
depends on the distance traveled and the frequency of the GWs.

% Effect on inclination PE
In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for a nonprecessing binary.
This modification changes the ratio of the left-handed and right-handed polarization of the waveform, which affects the PE on the inclination as well.
The relationship between the amplitude ratio and the inclination is
\begin{equation}
    \frac{h_\mathrm{L}}{h_\mathrm{R}}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,,
\end{equation}
where $\iota$ is the inclination, the angle between our line of sight and the orbital angular
momentum of the binary. Therefore, the PE on $\iota$ depends on the amplitude ratio.

@@

In \citet{Okounkova_2022}, they assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
This would create a degeneracy between $\kappa$ and $\iota$, as they can affect the amplitude ratio in the same way.
To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
As a result, many pairs of $\iota$ and $\kappa$ can be plausible with the frequency independence.
